---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #3 - instructions

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}
library(tidyverse)
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)

#gavins<-tally(group_by(av,Death1,Death2,Death3,Death4,Death5, sort = FALSE))

deaths <- av %>% pivot_longer(cols = starts_with("Death"), names_to = "Time", values_to = "Death")
head(deaths)
#deaths$Time <- parse_number(deaths$Time)
#head(deaths)

gavin4<-deaths%>%group_by(Name.Alias, URL)%>%count(Death)

gavin5 = filter(gavin4, Death == "YES")

  


```

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`. 

Similarly, deal with the returns of characters.

```{r}
#doing the same thing we did for deaths with returns now

returns <- av %>% pivot_longer(cols = starts_with("Return"), names_to = "Time2", values_to = "return")
head(returns)
#deaths$Time <- parse_number(deaths$Time)
#head(deaths)

gavin6<-returns%>%group_by(Name.Alias, URL)%>%count(return)

returns = filter(gavin6, return == "YES")



# merging our two dataframes into one

combined<- merge(gavin5,gavin6, by=c("Name.Alias", "URL"))


```

Based on these datasets calculate the average number of deaths an Avenger suffers. 

```{r}

deaths2 <- filter(deaths, deaths$Death == "YES")
deaths2$Time <- parse_number(deaths2$Time)
head(deaths2)
mean(deaths2$Time, na.rm = TRUE)
```
**The mean number of deaths an Avenger suffers is 1.3.**

## Individually

For each team member, copy this part of the report. 

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement

> Quote the statement you are planning to fact-check.

### Include the code

Make sure to include the code to derive the (numeric) fact for the statement

### Include your answer

Include at least one sentence discussing the result of your fact-checking endeavor.

Upload your changes to the repository. Discuss and refine answers as a team.

## Individually - Emily Maruska

For each team member, copy this part of the report. 

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement

> "There’s a 2-in-3 chance that a member of the Avengers returned from their first stint in the afterlife"

### Include the code

Make sure to include the code to derive the (numeric) fact for the statement
```{r}
dead <- av %>% pivot_longer(cols = starts_with("Death"), names_to = "Time", values_to = "Death")
dead$Time <- parse_number(dead$Time)

numReturned <- dead %>% filter(dead$Return1 == "YES")
nrow(numReturned)

totalDied <- dead %>% filter(dead$Return1 != "")
nrow(totalDied)

chanceToReturnFirstTime <- nrow(numReturned) / nrow(totalDied)
chanceToReturnFirstTime
```

### Include your answer

**The data in the article is really accurate considering I counted all the avengers that returned the first time, then counted all the avengers regardless of if they returned or died and took the number returned over the total died and got 0.6666 which translates directly to a "2-in-3 chance"**


## Individually - Gavin Anderson

For each team member, copy this part of the report. 

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement

> "Of the nine Avengers we see on screen — Iron Man, Hulk, Captain America, Thor, Hawkeye, Black Widow, Scarlet Witch, Quicksilver and The Vision — every single one of them has died at least once in the course of their time Avenging in the comics. In fact, Hawkeye died twice!"

### Include the code

```{r}
combined2<-combined


#made a dataframe with only mcu avengers included
MCU_only<-av[c(3,4,5,7,8,9,16,10,14),-c(4:10)]

head(MCU_only)

#and verified that every single current mcu character died at least once 

```

### Include your answer

Looking at the data frame "MCU_only" we can see that all of the onscreen Avengers have indeed died at least once.


## Individually - Carolyn Jones

### FiveThirtyEight Statement

> "Out of 173 listed Avengers, my analysis found that 69 had died at least one time after they joined the team. That’s about 40 percent"

### Include the code

```{r}
deaths$Time <- parse_number(deaths$Time)
deaths <- filter(deaths, deaths$Death == "YES" & deaths$Time == 1)
nrow(deaths) # 69

69/173 # percentage of deaths
```


### Include your answer

After filtering the deaths dataset to only contain avengers who died at least 1 time, the number of remaining avengers was 69, which is accurate to what the article said. I also divided this number by the total number of avengers to confirm that it was about 40%

## Individually - John Nesnidal

For each team member, copy this part of the report. 

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement

> "Given the Avengers’ 53 years in operation and overall mortality rate, fans of the comics can expect one current or former member to die every seven months or so, with a permanent death occurring once every 20 months."

### Include the code

Make sure to include the code to derive the (numeric) fact for the statement
```{r}


totalDeaths <- nrow(deaths2)
print(totalDeaths)

monthsPerDeath <- (53 * 12) / totalDeaths
print(monthsPerDeath)

# Months per death: 7.1

deaths3 <- deaths2 %>%
  filter(Return1 == "NO" | Return2 == "NO" | Return3 == "NO" |
           Return4 == "NO" | Return5 == "NO") %>%
  group_by(URL) %>%
  summarise(
    numDeaths = max(Time)
  )

totalPermDeaths <- nrow(deaths3)
print(totalPermDeaths)

monthsPerPermDeath <- (53 * 12) / totalPermDeaths
print(monthsPerPermDeath)

# Months per permanent death: 19.9


```

### Include your answer

**The data in the article is accurate, rounded to the nearest whole number. For the first statement (one death every 7 months), I took 53 * 12 and divided it by the number of rows in the deaths table. For the second statement (one permanent death every 20 months), I took 53 * 12 and divided it by the number of individual Avengers who had a NO reported in any of the return columns. These numbers come out to 7.1 months per death and 19.9 months per permanent death.**


